---
title: "Prediction Assignment Writeup"
author: "Raka Indra Rahmawan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This report aims to predict the manner in which exercises were performed based on sensor data. The target variable is `classe`.

## Data Loading and Preprocessing

```{r}
library(caret)
library(ggplot2)
library(dplyr)

# Load Data
train_data <- read.csv("pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
test_data <- read.csv("pml-testing.csv", na.strings = c("NA", "", "#DIV/0!"))

# Remove near-zero variance predictors
nzv <- nearZeroVar(train_data, saveMetrics=TRUE)
train_data <- train_data[, !nzv$nzv]

# Remove columns with too many missing values
train_data <- train_data[, colSums(is.na(train_data)) < nrow(train_data) * 0.6]

# Remove identifier columns
train_data <- train_data[, -c(1:6)]

# Ensure the target variable 'classe' is not removed
classe_column <- "classe"
common_columns <- intersect(colnames(train_data), colnames(test_data))
common_columns <- unique(c(common_columns, classe_column)) # Ensure 'classe' is retained

train_data <- train_data[, common_columns]
test_data <- test_data[, setdiff(common_columns, classe_column)]  # Test data has no 'classe'
```

## Model Training and Cross-Validation

```{r}
set.seed(123)
train_index <- createDataPartition(train_data$classe, p = 0.7, list = FALSE)
train_set <- train_data[train_index, ]
test_set <- train_data[-train_index, ]

model <- train(classe ~ ., data = train_set, method = "rf", trControl = trainControl(method = "cv", number = 5))

print(model)
```

## Model Evaluation

```{r}
predictions <- predict(model, test_set)

# Check for NA values in predictions
if (any(is.na(predictions))) {
  most_frequent_class <- names(sort(table(test_set$classe), decreasing = TRUE))[1]
  predictions[is.na(predictions)] <- most_frequent_class
}

# Ensure both predictions and test labels have common levels
common_levels <- unique(c(levels(factor(predictions)), levels(factor(test_set$classe))))
predictions <- factor(predictions, levels = common_levels)
test_set$classe <- factor(test_set$classe, levels = common_levels)

# Compute the confusion matrix
conf_matrix <- confusionMatrix(predictions, test_set$classe)
print(conf_matrix)
```

## Prediction on New Test Data

```{r}
test_predictions <- predict(model, test_data)

test_predictions
```

## Debugging

```{r}
# Check Unique Levels of Both Factors
predictions <- factor(predictions)
test_set$classe <- factor(test_set$classe)

str(predictions)
str(test_set$classe)

# Check distribution
table(predictions, useNA = "ifany")
table(test_set$classe, useNA = "ifany")
